Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-04-28 ms_pre_post.Rmd
Add Greta's comments up to Discussion

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values became substantially higher than in comparable simulations where $F_{a}$ values were fixed to zero and post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry to the degree that most females mated multiply given low costs of polyandry ($c_{P} < 0.005$; e.g., Figure 2A,B). Here, $P_{a}$ allele values increased from zero and persisted at low positive values (Figure 1B,D). Given higher costs of polyandry ($c_{P} \geq 0.005$), $P_{a}$ allele values still initially increased from zero, but then became slightly negative over generations (Figure 1F,H). Trajectories of allele values in individual simulations were typically highly stochastic, but were consistent in their long-term direction (Supporting Information Figures S1-S8). Overall, this implies that, even given strong inbreeding depression in offspring viability and simultaneous evolution of pre-copulatory and post-copulatory strategies, inbreeding avoidance might generally be a weak force driving polyandry evolution.

Strong post-copulatory inbreeding avoidance, manifested as very negative $F_{a}$ values, consistently evolved in all simulations where such evolution was allowed (black lines, Figure 1B,D,F,H). Evolution of post-copulatory inbreeding avoidance occurred even when $P_{a}$ values were expected to be slightly negative, and hence when there was selection against polyandry (Figure 1F,H). This reflects the threshold nature of phenotypic expression of polygenic polyandry, wherein random sampling of alleles means that polyandry is expressed by some females (i.e., $P_{g}>0$) even when mean $P_{a}$ values are negative (Figure 2). This means that, even in populations where female reproductive strategy evolves toward monandry, there is still commonly some opportunity for expression of post-copulatory inbreeding avoidance and associated selection that drives evolution of post-copulatory inbreeding avoidance.

@@

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When pre-copulatory but not post-copulatory inbreeding avoidance was costly ($c_{M}=0.02$ and $c_{F}=0$), pre-copulatory inbreeding avoidance initially evolved (i.e., negative $M_{a}$) but then evolved back towards random mating (i.e., $M_{a}=0$) following increasing evolution of post-copulatory inbreeding avoidance (Figure 3B). Similarly, when post-copulatory but not pre-copulatory inbreeding avoidance was costly ($c_{F}=0.02$ and $c_{M}=0$), post-copulatory inbreeding avoidance initially evolved ($F_{a}=0$) before evolving back to random fertilisation after ca $20000$ generations (Figure 3D).

@@

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. However, in this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding avoidance was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information Figure S10).

>>>>

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values became substantially higher than in comparable simulations where $F_{a}$ values were fixed to zero and post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry to the degree that most females mated multiply given low costs of polyandry ($c_{P} < 0.005$; e.g., Figure 2A,B). Here, $P_{a}$ allele values increased from zero and persisted at low positive values (Figure 1B,D). Given higher costs of polyandry ($c_{P} \geq 0.005$), $P_{a}$ allele values still initially increased from zero, but then became slightly negative over generations (Figure 1F,H). Trajectories of allele values in individual simulations were typically highly stochastic, but were consistent in their long-term direction (Supporting Information Figures S1-S8). Overall, post-copulatory inbreeding avoidance can drive evolution of polyandry as long as direct costs are sufficiently low, but high costs might cause such evolution of polyandry to be weak even given strong inbreeding depression in offspring viability.

Strong post-copulatory inbreeding avoidance, manifested as very negative $F_{a}$ values, consistently evolved in all simulations where such evolution was allowed (black lines, Figure 1B,D,F,H). Evolution of post-copulatory inbreeding avoidance occurred even when $P_{a}$ values were expected to be slightly negative, and hence when there was selection against alleles underlying polyandry (Figure 1F,H). This reflects the threshold nature of phenotypic expression of polygenic polyandry, wherein random sampling of alleles means that polyandry is expressed by some females (i.e., $P_{g}>0$) even when mean $P_{a}$ values are negative (Figure 2). This means that, even in populations where female reproductive strategy evolves toward monandry, there is still commonly some opportunity for expression of post-copulatory inbreeding avoidance and associated selection that drives evolution of post-copulatory inbreeding avoidance.

@@

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When pre-copulatory but not post-copulatory inbreeding avoidance was costly ($c_{M}=0.02$ and $c_{F}=0$), pre-copulatory inbreeding avoidance initially evolved (i.e., negative $M_{a}$) but then evolved back towards random mating (i.e., $M_{a}=0$) following increasing evolution of post-copulatory inbreeding avoidance and polyandry (Figure 3B). Similarly, when post-copulatory but not pre-copulatory inbreeding avoidance was costly ($c_{F}=0.02$ and $c_{M}=0$), post-copulatory inbreeding avoidance initially evolved ($F_{a}=0$) before evolving back to random fertilisation after ca $20000$ generations (Figure 3D).

@@

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. In this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding avoidance was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information Figure S10).
